\documentclass[defaultstyle,11pt]{thesis}

\usepackage{amssymb}		% to get all AMS symbols
\usepackage{graphicx}		% to insert figures
\usepackage{hyperref}		% PDF hyperreferences??

% self packages
\usepackage{multirow}
\usepackage{url}
\usepackage{amsmath}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage{enumitem}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator*{\argmin}{argmin}

\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}

%%%%%%%%%%%%   All the preamble material:   %%%%%%%%%%%%

\title{Metadata Matters: Adaptation Methods for Robust Document Classification}

\author{Xiaolei}{Huang}

\otherdegrees{B.E., Central China Normal University, 2013 \\
	      M.S., Chinese Academy of Sciences, 2016 \\
	      M.S., University of Colorado Boulder, 2019}

\degree{Doctor of Philosophy}		%  #1 {long descr.}
	{Ph.D., Information Science}		%  #2 {short descr.}

\dept{Department of}			%  #1 {designation}
	{Information Science}		%  #2 {name}

\advisor{Prof.}				%  #1 {title}
	{Robin Burke}			%  #2 {name}

\reader{Dr. Michael J. Paul}		%  2nd person to sign thesis
\readerThree{Prof. James H. Martin}		%  3rd person to sign thesis
\readerFour{Prof. Mark Dredze}		%  4th person to sign thesis
\readerFive{Prof. Byron Wallace}		%  5th person to sign thesis


\abstract{
Metadata, implicitly embedded in documents such as time, demographic factors and user interests, can cause language variations and impact performance of document classifiers. For example, language shifts over periods of time, and males and females express sentiment differently. However, models for document classification, the automatic categorization of documents into categories, typically ignore document metadata. 
In this thesis, we focus on two types of document metadata, temporality and user factors.
We propose to use domain adaptation by treating each metadata attribute as domains (e.g., gender domains: male vs. female), aiming to integrate temporality and user factors into document classifiers and improve classification performance.

First, we propose \textit{temporality adaptation} that explicitly incorporates time into the representation learning process via feature augmentation and diachronic word embedding. The feature augmentation method aims to learn time-independent feature weights for document classifiers. We then develop an end-to-end time-adapted model with the diachronic word embedding under a time-driven framework.
Second, we propose \textit{user factor adaptation} that models demographic attributes and user interests using multitask learning.
To model demographic attributes, document classifiers jointly predict demographic factors and document categories.
We further develop a multitask user embedding that jointly learns language, user behaviors and user interests.
We examine and visualize impacts of temporality and user factor on word, topic, semantic and classifier levels.

Benefits of adapting demographic attributes motivate us to examine if domain adaptation can reduce demographic biases. We release a multilingual hate speech corpus with author-level demographic labels. We examine demographic variations of user language and demographic biases of document classifiers. Following this, to reduce demographic bias, we apply a feature augmentation method to learn demographic-independent classifiers.
}

\dedication[Dedication]{	% NEVER use \OnePageChapter here.
	For my family and friends, who have supported and shaped me on this academic journey.
	}

\acknowledgements{	%\OnePageChapter	% *MUST* BE ONLY ONE PAGE!
    This thesis can not be done with supports from my thesis committee members and NSF IIS-1657338.
    I want to give my highest gratitude to my Ph.D. advisor, Michael J. Paul.
    He is a fantastic advisor who provides me with his generous support and endless help but not pressure or constraint along my four years of enjoyable and delightful study.
    Michael sets me an awesome example as a nice and good advisor that I could follow in the future.
    I really appreciate Mark Dredze's guidance and support on my research direction and career development. 
    Mark involved me in several projects through past years and gave me a really valuable visiting chance at the Johns Hopkins University. 
    His technical depth, application broadness as well as balance between work and life is a life-long inspiration for me.
    I want to thank Robin Burke for his feedback and help on my research and career development. 
    I am really inspired by his research and work experience from industry to academia.
    I had the fortune to have James H. Martin as my thesis committee member.
    Jim inspired me in many ways in both research and education that rebuilt my confidence when I started my doctoral study. 
    I first met Byron Wallace when I applied for the doctoral study at UT-Austin.
    I am glad and lucky to have him in my thesis committee.
    He made significant contributions in this thesis writing.
    His passion in research and health-related applications has deeply influenced my career goals and research agenda.
    
    Beyond my thesis committee, I was fortunate to have a number of collaborators from both academia and industry.
    I enjoyed insightful discussions with David Broniatowski, Amelia Jamison, Sandra C. Quinn on multiple public health projects. 
    It was really lucky to have my first industrial internship at Adobe Research with Franck Dernoncourt and Walter Chang. 
    I really appreciate their generous support and continuous collaborations since then. 
    I am lucky to spend two summers at the USC, where I worked with Stefan Scherer, Jonathan May and Nanyun Peng. 
    I benefited a lot from helpful and encouraging discussions with them. 
    Nanyun sets a role model for me to pursue an academic career.
    It was an exciting experience at Amazon Lab 126 with Minmin Shen and Angeliki Metallinou. The unique project broadened my eyes and enhanced my skills of parallel GPUs.
    The summer internship can not be so enjoyable without my friends including but not limit to: Anthony Michael Colas, Logan Lebanoff, Zhuowan Li, Naoto Inoue, Lixing Liu, Zheng Song and Jason Wu. My apologies for any omissions.
    
    I was honored to get into the PhD program at CU Boulder. 
    I benefit from the collaboration and advice from faculty members over the years, including but not limit to Leysia Palen, Jed Brubaker, Stephen Voida, Jordan Boyd-Graber (now at UMD), Danielle Szafir, Martha Palmer, Amy Voida and Casey Fiesler. I also owe a lot to my peer colleagues whose important discussions have helped me along the way: Linzi Xing, Ben Niu, Ashlynn Daughton, Shudong Hao, Dianna Radpour, Yoshinari Fujinuma, Todd Amodeo, Vivian Lai, Zijiao Yang, Wei-Te Chen, Matthew Whitlock, Wendy Norris, Bo Cao, Caleb Hsu, Alvin Grissom II and Wei-Lun Chao.
    My research career started by the collaborations with David Chiu and Tingshao Zhu when I was in China. Their insightful discussions and helpful guidance initialized my career in NLP.
    I met Aimao Zhang when I was in Georgia Southern, and her story always encourages me along my life: never too late.
    
    Finally, I am grateful for the support from my family and loved ones over the years, especially my parents Huang S.M. and Zhang Q.L., who give me so much love and share me with their valuable experience. I am very fortunate to meet my American family in Georgia, Gail Johnson Davis and Ed Davis. Their love is warm and sweet. Yang T.T. lighted up my enthusiasm for life. We feel like we have known each other for many years while we just met each other a few months ago.
}

% \IRBprotocol{E927F29.001X}	% optional!

\ToCisShort	% use this only for 1-page Table of Contents

\LoFisShort	% use this only for 1-page Table of Figures
% \emptyLoF	% use this if there is no List of Figures

\LoTisShort	% use this only for 1-page Table of Tables
% \emptyLoT	% use this if there is no List of Tables

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%       BEGIN DOCUMENT...         %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\input{macros.tex}
\input{chapter1.tex}
\input{chapter2.tex}
\input{chapter3.tex}
\input{chapter4.tex}
\input{chapter5.tex}
\input{chapter6.tex}

%%%%%%%%%   then the Bibliography, if any   %%%%%%%%%
\bibliographystyle{acm}	% or "siam", or "alpha", etc.
\bibliography{refs}		% Bib database in "refs.bib"


%%%%%%%%%   then the Appendices, if any   %%%%%%%%%
% \appendix
% \input Vita.tex
% \input appendixB.tex


\end{document}
